---
title: "COVID_19_Project"
author: "Kashan Ahmed"
date: "12/7/2021"
output:
  
  html_document: default
  pdf_document: default
---


### Importing Data:

Library in tidyverse package
Importing Data Sets from Github repository of John Hopkins Website in url_in.
Set file_names and concatenate all four files.
Then string concatenate url_in and file_names in urls.
```{r}
library(tidyverse) 
library(lubridate)

url_in <- "https://raw.githubusercontent.com/CSSEGISandDATA/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/"

file_names <- c("time_series_covid19_confirmed_global.csv",
                "time_series_covid19_deaths_global.csv",
                "time_series_covid19_confirmed_US.csv",
                "time_series_covid19_deaths_US.csv")
urls <- str_c(url_in,file_names)
```

Read Data-Sets and putting them in variable names.

```{r}
global_cases <- read_csv(urls[1])
global_deaths <- read_csv(urls[2])
US_cases <- read_csv(urls[3])
US_deaths <- read_csv(urls[4])
```

### Tidying and Transforming Data

We will pivot longer all columns except Province/State, Country/Region, Lat, Long. Change names_to = "date", values_to = "cases" and remove Lat and Long. 
Join Global cases and Global deaths in Global.
```{r}
global_cases <- global_cases %>% 
  pivot_longer(cols = -c(`Province/State`,
                         `Country/Region`, Lat, Long),
              names_to = "date",
              values_to = "cases") %>% 
  select(-c(Lat, Long))
  
global_deaths <- global_deaths %>% 
  pivot_longer(cols = -c(`Province/State`,
                         `Country/Region`, Lat, Long),
              names_to = "date",
              values_to = "deaths") %>% 
  select(-c(Lat, Long))

global <- global_cases %>% 
    full_join(global_deaths) %>% 
    rename(Country_Region =`Country/Region`,
           Province_State = `Province/State`) %>% 
    mutate(date = mdy(date))

global

summary(global)
```
Filter out Zero cases
```{r}
global <- global %>%  filter(cases > 0)
summary(global)

```
Filter out cases more than 28000000
```{r}
global %>% filter(cases > 28000000) 
```
Now we will work with US cases and do the same pivoting and joing deaths and cases in US.
```{r warning=FALSE}
US_cases %>% 
  pivot_longer(cols = -(UID:Combined_Key),
              names_to = "date",
              values_to = "cases")
              
US_cases <- US_cases %>% 
  pivot_longer(cols = -(UID:Combined_Key),
              names_to = "date",
              values_to = "cases") %>% 
  select(Admin2:cases) %>% 
  mutate(date = mdy(date)) %>% 
  select(-c(Lat, Long_))
  
US_deaths <- US_deaths %>% 
  pivot_longer(cols = -(UID:Population),
              names_to = "date",
              values_to = "deaths") %>% 
  select(Admin2:deaths) %>% 
  mutate(date = mdy(date)) %>% 
  select(-c(Lat, Long_))
  
US <- US_cases %>% 
    full_join(US_deaths)
```

further wrangling global
```{r}
global <- global %>% 
  unite("Combined_Key",
        c(Province_State, Country_Region),
        sep = ",",
        na.rm = TRUE,
        remove = FALSE)
global
```

Adding Population Data
```{r}
uid_lookup_url <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/UID_ISO_FIPS_LookUp_Table.csv"
uid <- read_csv(uid_lookup_url) %>% 
    select(-c(Lat, Long_, Combined_Key, code3, iso2, iso3, Admin2))
```

```{r}
global <-global %>% 
  left_join(uid, by = c("Province_State", "Country_Region")) %>% 
  select(-c(UID, FIPS)) %>% 
  select (Province_State, Country_Region, date,
          cases, deaths, Population,
          Combined_Key)
global
```

### Analysing and Visualizing Data

```{r warning=FALSE}
US_by_state <- US %>% 
  group_by(Province_State, Country_Region, date) %>% 
  summarize(cases = sum(cases), deaths = sum(deaths),
            Population = sum(Population)) %>% 
  mutate(deaths_per_mill = deaths * 1000000/Population) %>% 
  select(Province_State, Country_Region, date,
         cases, deaths, deaths_per_mill, Population) %>% 
  ungroup()
  
US_totals <- US_by_state %>%
    group_by(Country_Region, date) %>% 
   summarize (cases = sum(cases), deaths = sum(deaths),
              Population = sum(Population)) %>% 
   mutate(deaths_per_mill = deaths * 1000000/Population) %>% 
  select(Country_Region, date,
         cases, deaths, deaths_per_mill, Population) %>% 
  ungroup()
  
US_totals

```

Covid-19 in US

```{r}
US_totals %>% 
  filter(cases>0) %>% 
  ggplot(aes(x = date, y = cases))+
  geom_line(aes(color = "cases"))+
  geom_point(aes(color = "cases"))+
  geom_line(aes(y = deaths, color = "deaths"))+
  geom_point(aes(y = deaths, color = "deaths"))+
  scale_y_log10() +
  theme(legend.position = "bottom", 
        axis.text.x = element_text(angle = 90))+
  labs(title = "COVID19 in US", y = NULL)
```

Covid-19 in Virginia

```{r warning=FALSE}
state <- "Virginia"
US_by_state %>%
  filter(Province_State == state) %>% 
  filter(cases > 0) %>% 
  ggplot(aes(x = date, y = cases))+
  geom_line(aes(color = "cases"))+
  geom_point(aes(color = "cases"))+
  geom_line(aes(y = deaths, color = "deaths"))+
  geom_point(aes(y=deaths, color = "deaths"))+
  scale_y_log10()+
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90))+
  labs(title = str_c("COVID19 in ",state), y = NULL)
```

Covid-19 in Pakistan


```{r warning=FALSE}
global_totals<- global %>%
    group_by(Country_Region, date) %>% 
   summarize (cases = sum(cases), deaths = sum(deaths),
              Population = sum(Population)) %>% 
   mutate(deaths_per_mill = deaths * 1000000/Population) %>% 
  select(Country_Region, date,
         cases, deaths, deaths_per_mill, Population) %>% 
  ungroup()

global_totals

Global_by_country <- global %>% 
    mutate(new_cases = cases - lag(cases),
          new_deaths = deaths - lag(deaths))
global_totals <- global_totals %>% 
  mutate(new_cases = cases - lag(cases),
          new_deaths = deaths - lag(deaths))

country <- "Pakistan"
Global_by_country %>%
  filter(Country_Region == country) %>% 
  filter(cases > 0) %>% 
  ggplot(aes(x = date, y = new_cases))+
  geom_line(aes(color = "new_cases"))+
  geom_point(aes(color = "new_cases"))+
  geom_line(aes(y = new_deaths, color = "new_deaths"))+
  geom_point(aes(y= new_deaths, color = "new_deaths"))+
  scale_y_log10()+
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90))+
  labs(title = str_c("COVID19 in ",country), y = NULL)

          

```




New Cases in US
```{r}
US_by_state <- US_by_state %>% 
    mutate(new_cases = cases - lag(cases),
          new_deaths = deaths - lag(deaths))
US_totals <- US_totals %>% 
  mutate(new_cases = cases - lag(cases),
          new_deaths = deaths - lag(deaths))

tail(US_totals %>%  select(new_cases, new_deaths, everything()))

US_totals %>%
  ggplot(aes(x = date, y = cases))+
  geom_line(aes(color = "new_cases"))+
  geom_point(aes(color = "new_cases"))+
  geom_line(aes(y = new_deaths, color = "new_deaths"))+
  geom_point(aes(y = new_deaths, color = "new_deaths"))+
  scale_y_log10()+
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90))+
  labs(title = ("COVID19 in US"), y = NULL)
```


new case in Virginia

```{r warning=FALSE}
state <- "Virginia"
US_by_state %>%
  filter(Province_State == state) %>% 
  filter(cases > 0) %>% 
  ggplot(aes(x = date, y = new_cases))+
  geom_line(aes(color = "new_cases"))+
  geom_point(aes(color = "new_cases"))+
  geom_line(aes(y = new_deaths, color = "new_deaths"))+
  geom_point(aes(y= new_deaths, color = "new_deaths"))+
  scale_y_log10()+
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90))+
  labs(title = str_c("COVID19 in ",state), y = NULL)
```


Global cases Covid-19 cases

```{r warning=FALSE}
global_totals<- global %>%
    group_by(Country_Region, date) %>% 
   summarize (cases = sum(cases), deaths = sum(deaths),
              Population = sum(Population)) %>% 
   mutate(deaths_per_mill = deaths * 1000000/Population) %>% 
  select(Country_Region, date,
         cases, deaths, deaths_per_mill, Population) %>% 
  ungroup()

global_totals

 global_totals%>% 
  filter(cases>0) %>% 
  ggplot(aes(x = date, y = cases))+
  geom_line(aes(color = "cases"))+
  geom_point(aes(color = "cases"))+
  geom_line(aes(y = deaths, color = "deaths"))+
  geom_point(aes(y = deaths, color = "deaths"))+
  scale_y_log10() +
  theme(legend.position = "bottom", 
        axis.text.x = element_text(angle = 90))+
  labs(title = "COVID19 around the Globe", y = NULL)
 
```


### Analysing and Modeling Data

```{r}
US_state_totals <- US_by_state %>% 
  group_by(Province_State) %>% 
  summarize(deaths = max(deaths), cases = max(cases),
            population = max(Population),
            cases_per_thou = 1000*cases/population,
            deaths_per_thou = 1000 * deaths/population) %>% 
  filter(cases > 0, population > 0)

US_state_totals %>% 
  slice_min(deaths_per_thou, n = 10) %>% 
select(deaths_per_thou,cases_per_thou,everything())

US_state_totals %>% 
  slice_max(deaths_per_thou, n = 10) %>% 
select(deaths_per_thou,cases_per_thou,everything())
```
modeling

```{r}
mod <- lm(deaths_per_thou ~ cases_per_thou, data = US_state_totals)
summary(mod)
```

```{r}
US_state_totals %>% mutate(pred = predict(mod))
US_tot_w_pred <- US_state_totals %>% mutate (pred = predict(mod))
US_tot_w_pred

US_tot_w_pred %>% ggplot() +
  geom_point(aes(x = cases_per_thou, y = deaths_per_thou), color = "red3")+
   geom_line(aes(x = cases_per_thou, y = pred), color = "green")

```

Global Data Modeled

```{r}
Global_bycountry_totals <- Global_by_country %>% 
  group_by(Country_Region) %>% 
  summarize(deaths = max(deaths), cases = max(cases),
            population = max(Population),
            cases_per_thou = 1000*cases/population,
            deaths_per_thou = 1000 * deaths/population) %>% 
  filter(cases > 0, population > 0)

Global_bycountry_totals %>% 
  slice_min(deaths_per_thou, n = 10) %>% 
  select(deaths_per_thou,cases_per_thou,everything())

mod2 <- lm(deaths_per_thou ~ cases_per_thou, data = Global_bycountry_totals)
summary(mod2)

Global_bycountry_totals%>%  slice_max(cases_per_thou)
Global_bycountry_totals%>% mutate(pred = predict(mod2))

Global_tot_w_pred <- Global_bycountry_totals %>% mutate (pred = predict(mod2))
Global_tot_w_pred

Global_tot_w_pred %>% ggplot() +
  geom_point(aes(x = cases_per_thou, y = deaths_per_thou), color = "red3")+
  geom_line(aes(x = cases_per_thou, y = pred), color = "green")
```


### Conclusion

I have analysed Covid-19 data and visualized according to US and Globally, I have visualized total cases and new cases for US and Virginia and total cases for Global level and Pakistan. There are two linear models in this analysis, one for US and other for Global.
I tried to analysed the data to the best of my abilities but I am sure there must be some bias in my analysis. As I am interested in US state of Virginia and Country Pakistan, I have particularly visualized those two places.

The Covid-19 pandemic is still on going, so the data and analysis will happen for years to come.
